---
title: "STAT 345 Midterm Project"
author: "Ryan Winder"
font: 12pt
date: "Due March 31"
output:
  html_document: default
  word_document: default
  pdf_document: default
---

```{r, setup, include=FALSE}
knitr::opts_chunk$set(out.width = '80%', collapse = TRUE, warning=FALSE, message=FALSE)
library(tidyverse)
library(png)
```

```{r message=FALSE, warning=FALSE, include = FALSE}
# nothing special going on here, loading in the data and the necessary libraries
devtools::install_github("abresler/nbastatR")
library(nbastatR)
library(ggplot2)
library(ggthemes)
library(gganimate)
library(dplyr)
Sys.setenv("VROOM_CONNECTION_SIZE" = 131072 * 2) # setting the correct connection
```

```{r warning=FALSE, include=FALSE}
# Will be looking at seasons for the Golden State warriors
GSW_shots_year_2000 <- teams_shots(teams = "Golden State Warriors", seasons = 2000)
GSW_shots_year_2005 <- teams_shots(teams = "Golden State Warriors", seasons = 2005)
GSW_shots_year_2020 <- teams_shots(teams = "Golden State Warriors", seasons = 2020)
GSW_shots_year_2022 <- teams_shots(teams = "Golden State Warriors", seasons = 2022)
# combine the years to look at their change over time
GSW_shots_combined_years <- rbind(GSW_shots_year_2000, GSW_shots_year_2005, GSW_shots_year_2020, GSW_shots_year_2022)
# combine the first year we are looking at and the most recent year
GSW_shots_change <- rbind(GSW_shots_year_2000,GSW_shots_year_2022)
```

```{r, warning=FALSE, include=FALSE}
### For reference, look back to the first plot above
# Understanding the plot of shots for one season of the Golden State Warriors
# Golden State's blue is #1D428A 
# Golden State's yellow is #FFC72C
One_Season <- ggplot(data=GSW_shots_year_2000, aes(x = locationX, y=locationY - 500)) + # needed to bring the points down to were we can see them
  geom_point(alpha = 0.4, aes(color = GSW_shots_year_2000$typeEvent)) +  # changing opacity and color
  scale_color_manual(values = c("#1D428A","#FFC72C")) + # setting the colors to Golden State's colors
  labs(x = "", y = "", color = "Type of Shot") +
  xlim(-300,300) + 
  ylim(-550,0) + 
  ggtitle("GSW Shots without court during 2000")
```

```{r, include = FALSE}
# source in a R file for obtaining a court object
source("CourtforMidterm.R")
Court <- courtplot() # setting the court object to court
```

```{r, warning=FALSE, include=FALSE}
### For reference, look back to the animation in the third plot above
# add to the court plot
Animated_Seasons <- Court +
  geom_point(data = GSW_shots_combined_years, aes(x = locationX, y = locationY, color = typeEvent), alpha = 0.25) + # set data to the comnbined years of interest, and color the points to makes/misses
  scale_color_manual(values = c("#1D428A","#FFC72C")) + # set the colors of the points to the colors of Golden State
  labs(x = "", y = "", title = 'Missed & Made Shots during {closest_state}',color = "Type of Shot") + # make the x and y have no title, set the title of the plot to update depending on the year we are looking at, and update the title of the legend 
  transition_states(yearSeason, transition_length = 0.25, state_length = 3) + # transition based on the year of the season, and view the season for 3 seconds, and take 0.25 seconds to switch to the next season
  enter_fade() + # fade when the shots for the season show up
  exit_shrink() + # shrink when the shots for the season go away
  theme_classic() # set the theme to classic
```

```{r, include=FALSE}
### For reference, look back to the fourth plot of four plots
# add to the court plot again
Shooting_Difference <- Court +
  geom_point(data = GSW_shots_change, aes(x = locationX, y = locationY, color = typeEvent), alpha = 0.15) + # set the data to the shots change, the color to miss or make, and the opacity
  scale_color_manual(values = c("#1D428A","#FFC72C")) + # set the colors for misses/makes to the Golden State colors
  labs(x = "", y = "", title = "Shooting Change Over the Years",color = "Type of Shot") + # make sure the x and y axis have no title, set the overarching title, and set the legend title
  facet_wrap(yearSeason~typeEvent, nrow = 2) + # make 4 graphs, based on the season(which there are two), and the type of shot(which is miss or make), and have two rows
  theme_classic() # set theme to classic
```

```{r, warning=FALSE, include=FALSE}
FT_Wins <- teams_annual_stats(teams = "Golden State Warriors") %>%
  filter(teamName == "Warriors") %>%
  select(pctWins,pctFT,isConferenceChampion)
```

```{r}
One_Season
```
<br>
- Made shots are the blue points <br>
- Missed shots are the yellow points <br>
- The outline of the three point line can be visualized <br>
- The greater the density in a location depicts that a greater number of shots were taken there <br>
--- The greatest density for this season appears to be along the baseline <br>
<br>
```{r}
Court
```
<br>
- This court will be great for visualizing what we are looking at <br>
- All the dimensions should be alligned to the shooting data <br>
<br>
```{r, warning=FALSE}
Animated_Seasons
```
<br>
- The shooting patterns appear to change season to season <br>
- No huge difference from 2000 to 2005 <br>
- Big difference from 2005 to 2020 <br>
--- A far greater density of shots can be seen in season 2020 around the three point line <br>
--- Far fewer shots taken inside the three point line and not near the basket <br>
------ Both of the two notes prior are even more so noticeable in season 2022 <br>
- Over time, the three point shooting has become more prominent <br>
<br>
```{r, warning=FALSE}
Shooting_Difference
```
<br>
- The densities are very similar for each plot of the same season <br>
--- No locations on the court indicate a far greater density from shots made compared to shots missed or vice versa <br>
------  Therefore, there does not appear to be any lucky or unlucky location on the court <br>
- All plots have a great density near the basket <br>
- Season 2000 has a far greater density of shots both missed and made inside the three point line <br>
- Season 2022 has a far greater density of shots both missed and made beyond the three point line <br>
<br> <br> <br> <br> <br>

<div align = "center"><font size="4">Further Reading</font></div>
<br>
- The data for this team goes further back, however looking at the changes in shots over a more recent period is of greater interest <br>
<br>
First Plot of Season 2000: <br>
- We can visualize the outline of the three point line from the shooting pattern on the first plot, which we can use that to better visualize by putting these points over a basketball court <br>
- The greatest density of shots is right where we would expect the basket to be <br>
--- The next areas of greatest density appears to be along the baseline, which is not what one would expect during the game today, indicating that we can expect to see large changes as we look into more recent seasons <br>
- There are a handful of shots far beyond the outlined three point line <br>
--- All of these except one are misses, indicating these are undesirable shots. <br>
- The proportion of missed(yellow) shots appears to be slightly more prominent that made(blue) shots, which we would expect, because the average field goal percentage of recent NBA seasons is roughly 45%. <br>
<br>
<br>
Animated Plot: <br>
- The areas of the greatest number of shots taken appears to change each season, other than a great density of shots taken right near the basket <br>
--- The seasons of 2000 and 2005 have a large number of shots scattered around inside the three point line that is not right near the basket <br>
- Season 2000 had a large majority of shots running along the baseline, which is nowhere present in the plots of seasons 2020 and 2022 <br>
- Season 2005 had the majority of shots right next to the basket and started to expand to taking many three point shots <br>
--- Seasons 2020 and 2022 had similar patterns with the greatest density of shots being right next to the basket and the majority of the remaining shots being three pointers <br>
------ Astonishingly fewer shots inside the three point line that is not near the basket compared to the seasons of 2000 and 2005. <br>
<br> <br>
Fourth plot of Shooting Change: <br>
- The change in shooting patterns is very noticeable from these plots <br>
- The density of shots inside the three point line that was not near the basket greatly transitioned to outside the three point line <br>
--- From this, we can see that the Golden State Warriors now views three pointers as more valuable shots, and when a three is not being taken, getting to the basket is the next best thing <br>
--- Also, not only are three pointers likely seen as more valuable by Golden State, the players have improved in shooting the three because in season 2000 only one shot far beyond the three point line was a make <br>
------ Whereas in the season 2022, there are multiple made shots far beyond the three point line with roughly the same number of missed shots far beyond the three point line as season 2000 <br>
- Goal for the players of Golden State <br>
--- Take open three pointers <br>
--- Try to get to the basket, if a shot is not open next to the basket, kick it back out for a three pointer <br>
<br> <br> <br>
<div align = "center"><font size="4">Extra Graphic</font></div> <br> <br>
```{r, warning=FALSE, include = FALSE}
avg_ft_avg <- mean(FT_Wins$pctFT) # the average of the averages of free throws for all seasons
avg_win_avg <- mean(FT_Wins$pctWins) # the average of the percentage of wins for all seasons
Golden_State_data <- ggplot(data=FT_Wins,aes(x = pctWins, y = pctFT)) + # assign new ggplot to a variable
  geom_point(aes(color = FT_Wins$isConferenceChampion)) + # color the points based on if conference champion
  scale_color_manual(values = c("#1D428A","#FFC72C")) + # set the colors to Golden State colors
  labs(x = "Percentage of Wins", y = "Free throw Percentage", title = "Golden States' Seasons since 1945", color = "Conference Winners") + # set the labels
  geom_smooth(method = "lm", formula = y ~ x, col = "white") + # add the regression line for this data, set the lines color to something easy to see
  geom_hline(yintercept = avg_ft_avg, linetype = "dashed", col = "white") + # add in vertical line for the average free throws
  geom_vline(xintercept = avg_win_avg, linetype = "dashed", col = "white") + # add in horizontal line for the average of wins for all seasons
  theme_dark() # set the theme to dark, because it looks good with Golden States colors
```

```{r}
Golden_State_data
```
<br>
- The Linear Regression line shows the relationship between the two variables <br>
--- Positive relationship, as Golden State has a higherer win percentage for a season, they have a higher free throw percentage, and vice versa <br>
- Vertical Line: The average of the average wins for all seasons <br>
- Horizontal Line: The average of the average of free throws for all seasons <br>
- As you can see <br>
--- There is only one occurrence of being conference winners both below the linear regression line, and below average of the free throw percentage line <br>
--- Therefore, a higher free throw percentage than the linear regression line during a season will make Golden State much more likely to be conference winners <br>
--- Every occurrence of being conference winners occurs when Golden State has a higher percentage of wins than the overall average of wins <br>
------Therefore, in order for Golden State to stand a better chance of being conference winners, they need to end the season with a higher win percentage than the average of their overall wins <br>
- This is the finding we would expect to see, however, it is interesting to see how large of an impact a high win percentage mixed with a high free throw percentage has on Golden State being the conference winners or not


